---
title: "Genomics of Inbreeding in recently bottlenecked Seychelles warblers"
author: "Kiran Gok Lune Lee"
date: "`r Sys.Date()`"
geometry: margin=2cm 
output:
  html_document:
    toc: yes
    toc_float:
      collapsed: no
    fig_width: 12
    fig_height: 10
    output_file: "Kiran_Lee_Genomics_Inbreeding_Seychelles_warblers.html"
  pdf_document:
    latex_engine: xelatex
    fig_width: 8
    fig_height: 6
    output_file: "Kiran_Lee_Genomics_Inbreeding_Seychelles_warblers.pdf"
---

## Set up environment

```{r setup, include=FALSE}
library(knitr)
knitr::opts_knit$set(root.dir = "~/Documents/GitHub/InProgressGenomicsInbreedingSeychellesWarblers/Data")
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

R.Version()
library(readr)
library(tidyverse)
library(dplyr)
library(readxl)
library(openxlsx)
library(ggplot2)
library("data.table") 
library("reshape2")
library(sequoia)
```

# Create a massive dataframe

## SeqIDs (1st round of imputation)

Load SeqIDs of sequenced samples after first round of imputation by STITCH.

```{r seqids1, eval = TRUE, echo = FALSE, include=TRUE}
CoverageExtraSamples <- read.delim("coveragefilenameallsortedextrasamples.txt",sep=" ",header=F)
names(CoverageExtraSamples)[names(CoverageExtraSamples) == 'V1'] <- 'SeqID'
names(CoverageExtraSamples)[names(CoverageExtraSamples) == 'V2'] <- 'Coverage'
CoverageExtraSamples$Filepath<-CoverageExtraSamples$SeqID
CoverageExtraSamples$Filepath <- gsub("^.{0,4}", "", CoverageExtraSamples$Filepath)
CoverageExtraSamples$SeqID<-sub('.', '', CoverageExtraSamples$SeqID)
CoverageExtraSamples<-CoverageExtraSamples %>% separate(SeqID, c('Filepath1', 'Filepath2', 'Filepath3', 'Filepath4','Plate','Filepath5','SeqID'), sep = '/', convert = TRUE)
CoverageExtraSamples = subset(CoverageExtraSamples, select = c(Plate, SeqID, Coverage, Filepath))
CoverageExtraSamples$ID<-CoverageExtraSamples$SeqID

```

## BirdIDs

Linking SeqIDs to BirdIDs as in field database

```{r birdids, eval= TRUE, warning=FALSE, include=TRUE}
##Files to match BirdIDs
Identifiers<-read_excel("SheffieldSubmissions.xlsx")
Identifiers26076<-read_excel("ID 26076_Sample information table.xlsx")
PilotIdentifiers<-read_excel("SamplesForPilotTargetCapture_290119_sortBTN_Qubit.xlsx")
LIMS26629renamed<-read.table("lims26629renamed.txt",sep=" ",header=F)
LIMS26757p1raw4<-read.csv("Samples for Sequencing 25072023.csv",colClasses=c("NULL",NA,NA,NA,NA,NA))
MissingLIMS26757p1raw4 <- data.frame(BirdID=c(6572,6145,6373,5904,6144,6651), 
                                    FieldRing=c(NA,NA,NA,NA,NA,NA),
                                    BTO=c(NA,NA,NA,NA,NA,NA), 
                                    BloodID=c(8178, 5883,6287,5620,5880,7239), 
                                    BloodTubeNumber=c(NA,NA,NA,NA,NA,NA), 
                                    stringsAsFactors=FALSE)
LIMS26757p1raw4<-rbind(LIMS26757p1raw4,MissingLIMS26757p1raw4)
BloodID<-read.csv("BloodID.csv")

##Clean ID numbers
LIMS26629renamed<-LIMS26629renamed %>% arrange(V1)
LIMS26629renamed$SeqID<-CoverageExtraSamples$SeqID[CoverageExtraSamples$Plate=='LIMS26629']
LIMS26629renamed<-LIMS26629renamed %>%
  mutate(SeqID=sort(SeqID))
LIMS26629renamed$V1<-sub(".*\\-", "", LIMS26629renamed$V1)
colnames(LIMS26629renamed)[colnames(LIMS26629renamed) == 'V1'] <- 'ID'

#Fix 6 samples that were mislabeled, as identified by Rowan
LIMS26629renamed$ID[which(LIMS26629renamed$SeqID == "100_ACAAGAACCT-CGATACTGAA_L002__all_mapped_rehead.bam")] <- 2544
LIMS26629renamed$ID[which(LIMS26629renamed$SeqID == "101_AGAGTATGTG-AGATGGCTTC_L002__all_mapped_rehead.bam")] <- 3227
LIMS26629renamed$ID[which(LIMS26629renamed$SeqID == "96_CAACCATACA-ACCGGTTATA_L002__all_mapped_rehead.bam")] <- 764
LIMS26629renamed$ID[which(LIMS26629renamed$SeqID == "97_GTAGGCCGTT-GCCACTGTCT_L002__all_mapped_rehead.bam")] <- 2688
LIMS26629renamed$ID[which(LIMS26629renamed$SeqID == "98_CGGATTGATC-AGTCACAACA_L002__all_mapped_rehead.bam")] <- 2493
LIMS26629renamed$ID[which(LIMS26629renamed$SeqID == "99_ACTGGCAAGA-TGTTGTCCAT_L002__all_mapped_rehead.bam")] <- 2522

CoverageExtraSamples$ID[CoverageExtraSamples$Plate=="LIMS26629"]<-NA
CoverageExtraSamples <- merge(CoverageExtraSamples,LIMS26629renamed,by="SeqID", all = TRUE) %>%
  mutate(ID = coalesce(ID.x, ID.y)) %>%
  select (-c(ID.x, ID.y))



PilotSequences<-subset(CoverageExtraSamples, CoverageExtraSamples$Plate=="LIMS24675"|CoverageExtraSamples$Plate=="LIMS25133")
PilotSequences$ID<-trimws(sapply(strsplit(PilotSequences$ID, "_"), `[[`, 2))
MissingBloodIDSequences<-subset(CoverageExtraSamples, CoverageExtraSamples$Plate == "LIMS26076p4"|CoverageExtraSamples$Plate == "LIMS26076raw")
MissingBloodIDSequences$ID<-as.numeric(sapply(strsplit(MissingBloodIDSequences$ID, "_"), "[[", 1))
BloodIDSequences<-subset(CoverageExtraSamples, CoverageExtraSamples$Plate!="LIMS24675"& CoverageExtraSamples$Plate!="LIMS25133"&CoverageExtraSamples$Plate != "LIMS26076p4"&CoverageExtraSamples$Plate != "LIMS26076raw" )
BloodIDSequences$ID <- sub('_repeat','',BloodIDSequences$ID)
BloodIDSequences$ID <- sub('-repeat','',BloodIDSequences$ID)
BloodIDSequences$ID<-trimws(sapply(strsplit(BloodIDSequences$ID, "_"), `[[`, 1))
BloodIDSequences$ID<-sub('.+-(.+)', '\\1', BloodIDSequences$ID)

##Pilot sequences use Blood Tube Number, the rest use BloodID
PilotSequences$Identifier<-paste("BloodTubeNumber")
MissingBloodIDSequences$Identifier<-paste("BloodID")
BloodIDSequences$Identifier<-paste("BloodID")

##Make IDs numeric
PilotSequences$ID<-as.numeric(PilotSequences$ID)
BloodIDSequences$ID<-as.numeric(BloodIDSequences$ID)
MissingBloodIDSequences$ID<-as.numeric(MissingBloodIDSequences$ID)
Identifiers26076$Sample_number<-as.numeric(Identifiers26076$`Sample number`)
Identifiers26076$BloodID<-as.numeric(Identifiers26076$`Sample name`)

#Join BirdIDs
PilotSequences <- PilotSequences %>% 
  left_join(select(PilotIdentifiers, BirdID, BloodTubeNumber), by = c("ID" = "BloodTubeNumber"))

BloodIDSequences$BloodID<-BloodIDSequences$ID
BloodIDSequences <- BloodIDSequences %>% 
  left_join(select(Identifiers, BirdID, BloodID), by = c("ID" = "BloodID"))
names(BloodIDSequences)[names(BloodIDSequences) == 'BirdID.x'] <- 'BirdID'
BloodIDSequences<-BloodIDSequences[!duplicated(BloodIDSequences), ]

MissingBloodIDSequences <- MissingBloodIDSequences %>% 
  left_join(select(Identifiers26076, Sample_number, BloodID), by = c("ID" = "Sample_number"))
MissingBloodIDSequences <- MissingBloodIDSequences %>% 
  left_join(select(Identifiers, BirdID, BloodID), by = c("BloodID" = "BloodID"))

##Concatenate into one file (2018 individuals including duplicates)
MissingBloodIDSequencesFormatted= subset(MissingBloodIDSequences, select = c(Plate, SeqID, Coverage, Filepath, BloodID, Identifier, BirdID))
colnames(MissingBloodIDSequencesFormatted)[colnames(MissingBloodIDSequencesFormatted) == 'BloodID'] <- 'ID'
BloodIDSequencesFormatted= subset(BloodIDSequences, select = c(Plate, SeqID, Coverage, Filepath, ID, Identifier, BirdID))
SequencedIndividualsBirdIDs=rbind(PilotSequences,MissingBloodIDSequencesFormatted,BloodIDSequencesFormatted)

##Add in missing BirdIDs from LIMS26757p1raw4
SequencedIndividualsBirdIDsExtra<- SequencedIndividualsBirdIDs %>% 
  left_join(select(LIMS26757p1raw4, BirdID, BloodID), by = c("ID" = "BloodID")) %>%
  mutate(BirdID = coalesce(BirdID.x, BirdID.y)) %>%
  select (-c(BirdID.x, BirdID.y))

SequencedIndividualsBirdIDsExtraDeduplicated<- SequencedIndividualsBirdIDsExtra %>% 
  group_by(BirdID) %>%
  top_n(1, abs(Coverage))
```

## Test panel

This script helped me find plates and well positions for cherry-picking samples at Liverpool to re-sequence. Samples were chosen following commands in the 9choosereferencepanel.sh bash script picking the best 250 coverage that were unrelated to at least a 2nd degree.

```{r testpanel, eval=TRUE, warning=FALSE }
##Choose samples to resequence
BestCoverageUnrelated<-read.table("unrelatedimputedchromosomes250sampled2nddegree.king.cutoff.in.id")
BestCoverageUnrelated$SeqID<-basename(BestCoverageUnrelated$V2)
BestCoverageUnrelatedSampleID<- merge(BestCoverageUnrelated, SequencedIndividualsBirdIDsExtraDeduplicated, by="SeqID", all = FALSE)
BestCoverageUnrelatedToResequence<-BestCoverageUnrelatedSampleID %>%
  select(SeqID, Plate, ID, Identifier, BirdID, Coverage )
###Add in plate and well numbers to help find libraries
BestCoverageUnrelatedToResequenceWellID<- merge(BestCoverageUnrelatedToResequence, Identifiers, by="BirdID", all = FALSE)
Identifiers26076<- Identifiers26076 %>% 
  left_join(select(BloodIDSequences, BloodID, BirdID) , by = c("BloodID" = "BloodID"))
BestCoverageUnrelatedToResequenceWellIDPlate <- merge(BestCoverageUnrelatedToResequenceWellID, Identifiers26076, by= "BirdID", all= TRUE)
BestCoverageUnrelatedToResequenceWellIDPlate <- subset(BestCoverageUnrelatedToResequenceWellIDPlate, !is.na(BestCoverageUnrelatedToResequenceWellIDPlate$SeqID))
BestCoverageUnrelatedToResequenceWellIDPlate <- BestCoverageUnrelatedToResequenceWellIDPlate[-c(18, 20, 21), ]

BestCoverageUnrelatedToResequenceWellIDPlate <- subset(BestCoverageUnrelatedToResequenceWellIDPlate,BestCoverageUnrelatedToResequenceWellIDPlate$Identifier=="BloodID")

BestCoverageUnrelatedToResequenceWellIDPlate <- subset(BestCoverageUnrelatedToResequenceWellIDPlate,BestCoverageUnrelatedToResequenceWellIDPlate$Plate.x!="LIMS24675")
BestCoverageUnrelatedToResequenceWellIDPlate <- subset(BestCoverageUnrelatedToResequenceWellIDPlate,BestCoverageUnrelatedToResequenceWellIDPlate$Plate.x!="LIMS25133")

openxlsx::write.xlsx(BestCoverageUnrelatedToResequenceWellIDPlate, file ="BestCoverageUnrelatedToResequenceWellIDPlate.xlsx", quote=FALSE)
getwd()

```

This script creates a set of UNIX commands to merge re-sequenced samples creating a higher coverage test set to assess imputation accuracy. Re-sequenced samples were verified as duplicateds using PLINK.

```{r warning=FALSE, include=FALSE}
##Merge reference panel
SequencedIndividualsBirdIDsExtraDeduplicated<- SequencedIndividualsBirdIDsExtra %>% 
  group_by(BirdID) %>%
  top_n(1, abs(Coverage))
###These samples have been checked they are indeed duplicates using PLINK
ReferencePanel<- read.delim("duplicatereferencepairs0.3na1.kin0",sep=" ",header=T,col.names=c("","Resequence","","SeqID","","","",""),row.names=NULL)
ReferencePanel<- subset(ReferencePanel, select = -c(X,X.1,X.2,X.3,X.4,X.5) )
ReferencePanel[] <- lapply(ReferencePanel, gsub, pattern='./Clean_aligned/', replacement='')

#Add in filepaths
ReferencePanel<-merge(ReferencePanel, SequencedIndividualsBirdIDsExtraDeduplicated , by="SeqID") %>%
                        select(Resequence,Filepath) 
ReferencePanel$Merged_output<-ReferencePanel$Filepath
ReferencePanel$Resequence <- sub("^", "/fastdata/bop21kgl/RawData/LIMS202103/Clean_aligned/", ReferencePanel$Resequence )
ReferencePanel$Merged_output<-gsub(".*/","",ReferencePanel$Merged_output)
ReferencePanel$Merged_output <- sub("^", "/fastdata/bop21kgl/RawData/LIMSMERGED/Clean_aligned/", ReferencePanel$Merged_output )
ReferencePanel$Command= "samtools merge"
ReferencePanel<-ReferencePanel[,c(4,3,1,2)]

#Write out commands for samtools to merge files
write.table(ReferencePanel, file = "ReferencePanel.txt", sep = " ",
            col.names = F, row.names = F, quote = FALSE)

openxlsx::write.xlsx(ReferencePanel, file ="ReferencePanel.xlsx", quote=FALSE)

```

## SeqIDs (2nd round of imputation with "test" panel)

```{r seqids2, eval=TRUE, echo=FALSE}

TestPanelSamples<- read.delim("coveragefilenameallsortedextrasamplesimputed.txt",sep=" ",header=F)
names(TestPanelSamples)[names(TestPanelSamples) == 'V1'] <- 'SeqID'
names(TestPanelSamples)[names(TestPanelSamples) == 'V2'] <- 'Coverage'
TestPanelSamples$Filepath<-TestPanelSamples$SeqID
TestPanelSamples$Filepath <- gsub("^.{0,4}", "", TestPanelSamples$Filepath)
TestPanelSamples$SeqID<-sub('.', '', TestPanelSamples$SeqID)
TestPanelSamples<-TestPanelSamples %>% separate(SeqID, c('Filepath1', 'Filepath2', 'Filepath3', 'Filepath4','Plate','Filepath5','SeqID'), sep = '/', convert = TRUE)
TestPanelSamples = subset(TestPanelSamples, select = c(Plate, SeqID, Coverage, Filepath))
TestPanelSamples$ID<-TestPanelSamples$SeqID
TestPanelSamples<-subset(TestPanelSamples, TestPanelSamples$Plate=="LIMSMERGED")

SequencedIndividualsBirdIDsExtraDeduplicated <- SequencedIndividualsBirdIDsExtraDeduplicated %>%
  left_join(TestPanelSamples %>% select(SeqID, Plate, Coverage, Filepath), by = "SeqID", suffix = c(".orig", ".new")) %>%
  mutate(
    Plate = coalesce(Plate.new, Plate.orig),  # Replace Plate.orig with Plate.new
    Coverage = coalesce(Coverage.new, Coverage.orig),  # Replace Coverage.orig with Coverage.new
    Filepath = coalesce(Filepath.new, Filepath.orig)  # Replace Coverage.orig with Coverage.new
  ) %>%
  select(-Plate.new, -Coverage.new, -Filepath.new)  # Remove the extra columns from the join

##Create downsample factor for samples to downsample using /SNPsSeychellesWarbler/blob/main/3_Imputation_accuracy/4downsamplecorrectcoverage.sh before imputing with STITCH
DownsampleReferencePanel<-subset(SequencedIndividualsBirdIDsExtraDeduplicated, SequencedIndividualsBirdIDsExtraDeduplicated$Plate=="LIMSMERGED")
DownsampleReferencePanel$Multiplier0.1X<-1/DownsampleReferencePanel$Coverage
DownsampleReferencePanel<-DownsampleReferencePanel %>%
  ungroup() %>%
  select(Filepath, Multiplier0.1X)

write.table(ReferencePanel, file = "downsamplescalingnew", sep = "\t",
            col.names = F, row.names = F, quote = FALSE)

```

## Genotype sex

This is important to check if we have assigned IDs correctly. If the sex of sample as calculated by imputed genotypes does not match the sex as recorded by microsatellites from the field database, something must be up.

```{r sex, eval=TRUE, echo = TRUE, tidy=FALSE}
##Load field database sex data
BirdIDSexYear <- read_excel("BirdIDSexYear.xlsx")


##Add in field database sex of bird
SequencedIndividualsBirdIDsExtraDeduplicated <- merge(SequencedIndividualsBirdIDsExtraDeduplicated, BirdIDSexYear[, c("BirdID", "Sex")], by = "BirdID", all.x = TRUE)
SequencedIndividualsBirdIDsExtraDeduplicated<-SequencedIndividualsBirdIDsExtraDeduplicated[!duplicated(SequencedIndividualsBirdIDsExtraDeduplicated),]

##Add in genotype sex of bird (heterozygosity of Z chromosome)
Zheterozygosity <- read.table("plink2.scount",header=TRUE, col.names=c("Filepath", "nHom", "nHet"))
Zheterozygosity$Het<- Zheterozygosity$nHet/Zheterozygosity$nHom
Zheterozygosity$Het <- round(Zheterozygosity$Het, 5)
hist(Zheterozygosity$Het, breaks=120)

##Add in genotype sex of bird (Z:W coverage and Autosome:Z coverage)
SexChromosomeCoverage<-read.table(file = "zw_autosomew_coverageall_clean.txt",sep=" ",header=F, col.names = c("SeqID","RatioZW", "RatioAutosomesW", "RatioAutosomesZ", "Mean_autosomes_cov", "Z_cov" , "W_cov"),  fill = TRUE)
SexChromosomeCoverage$Mean_autosomes_cov<-as.numeric(SexChromosomeCoverage$Mean_autosomes_cov)
SexChromosomeCoverage$RatioZW<-as.numeric(SexChromosomeCoverage$RatioZW)
SexChromosomeCoverage$RatioAutosomesW<-as.numeric(SexChromosomeCoverage$RatioAutosomesW)
SexChromosomeCoverage$RatioAutosomesZ<-as.numeric(SexChromosomeCoverage$RatioAutosomesZ)
hist(SexChromosomeCoverage$RatioZW, breaks=120)
hist(SexChromosomeCoverage$RatioAutosomesW, breaks=120)
hist(SexChromosomeCoverage$RatioAutosomesZ, breaks=120)
SexChromosomeCoverage<- SexChromosomeCoverage %>% 
  group_by(SeqID) %>%
  top_n(1, abs(Mean_autosomes_cov))
SexChromosomeCoverage<-SexChromosomeCoverage[!duplicated(SexChromosomeCoverage),]


#0.08 is dividing point. 0.076 to 0.084 is unclear.
Zheterozygosity <- Zheterozygosity %>%
  mutate(ZHeterozygositySex = case_when(
    Het > 0.084 ~ "0",           # If Het > 0.084, set Sex to "0"
    Het < 0.076 ~ "1",           # If Het < 0.076, set Sex to "1"
    #Het >= 0.076 & Het <= 0.084 ~ "Unsure",  # If Het is between 0.076 and 0.084, set Sex to "Unsure"
    TRUE ~ NA_character_         # In case any other values exist, handle them
  ))

SequencedIndividualsBirdIDsExtraDeduplicated <- SequencedIndividualsBirdIDsExtraDeduplicated %>%
  left_join(Zheterozygosity %>% select(Filepath, Het, ZHeterozygositySex), by = "Filepath")



ggplot(SequencedIndividualsBirdIDsExtraDeduplicated, aes(x = Het, fill = factor(Sex))) +
  geom_histogram(bins = 120, alpha = 0.7, position = "identity", color = "black") +
  scale_fill_manual(values = c("1" = "blue", "0" = "red")) +  # Blue for Sex=1, Red for Sex=0
  labs(x = "Het", y = "Frequency", title = "Histogram of Het with Sex-based Colors") +
  theme_minimal() +
  theme(legend.title = element_blank(), legend.position = "top")

#2 is dividing point. 1.99 to 2.01 is unclear.
SexChromosomeCoverage <- SexChromosomeCoverage %>%
  mutate(RatioAutosomesZSex = case_when(
    RatioAutosomesZ > 2.01 ~ "0",           # If Het > 2.01, set Sex to "0"
    RatioAutosomesZ < 1.99 ~ "1",           # If Het < 1.99, set Sex to "1"
    #RatioAutosomesZ >= 1.99 & RatioAutosomesZ <= 2.01 ~ "Unsure",  # If Het is between 1.99 and 2.01, set Sex to "Unsure"
    TRUE ~ NA_character_         # In case any other values exist, handle them
  ))

SequencedIndividualsBirdIDsExtraDeduplicated <- SequencedIndividualsBirdIDsExtraDeduplicated %>%
  left_join(SexChromosomeCoverage , by = "SeqID")



ggplot(SequencedIndividualsBirdIDsExtraDeduplicated, aes(x = RatioAutosomesZ, fill = factor(Sex))) +
  geom_histogram(bins = 120, alpha = 0.7, position = "identity", color = "black") +
  scale_fill_manual(values = c("1" = "blue", "0" = "red")) +  # Blue for Sex=1, Red for Sex=0
  labs(x = "RatioAutosomes:ZCoverage", y = "Frequency", title = "Histogram of Autosomes:Z Coverage Ratio with Sex-based Colors") +
  theme_minimal() +
  theme(legend.title = element_blank(), legend.position = "top")

SequencedIndividualsBirdIDsExtraDeduplicated <- SequencedIndividualsBirdIDsExtraDeduplicated %>%
  mutate(ZHeterozygositySex = case_when(
    Het > 0.080 ~ "0",           # If Het > 0.084, set Sex to "0"
    Het < 0.080 ~ "1",           # If Het < 0.076, set Sex to "1"
    #Het >= 0.076 & Het <= 0.084 ~ "Unsure",  # If Het is between 0.076 and 0.084, set Sex to "Unsure"
    TRUE ~ NA_character_         # In case any other values exist, handle them
  ))

SequencedIndividualsBirdIDsExtraDeduplicated <- SequencedIndividualsBirdIDsExtraDeduplicated %>%
  mutate(RatioAutosomesZSex = case_when(
    RatioAutosomesZ > 2.00 ~ "0",           # If Het > 2.01, set Sex to "0"
    RatioAutosomesZ < 2.00 ~ "1",           # If Het < 1.99, set Sex to "1"
    #RatioAutosomesZ >= 1.99 & RatioAutosomesZ <= 2.01 ~ "Unsure",  # If Het is between 1.99 and 2.01, set Sex to "Unsure"
    TRUE ~ NA_character_         # In case any other values exist, handle them
  ))

SequencedIndividualsBirdIDsExtraDeduplicated <- SequencedIndividualsBirdIDsExtraDeduplicated %>%
  mutate(ZHeterozygositySexMatch = ifelse(Sex == ZHeterozygositySex, "match", "no match"))

SequencedIndividualsBirdIDsExtraDeduplicated <- SequencedIndividualsBirdIDsExtraDeduplicated %>%
  mutate(RatioAutosomesZSexMatch = ifelse(Sex == RatioAutosomesZSex, "match", "no match"))

ggplot(SequencedIndividualsBirdIDsExtraDeduplicated, aes(x = Coverage, fill = ZHeterozygositySexMatch)) +
  geom_histogram(bins = 30, alpha = 0.7, position = "identity", color = "black") +
  scale_fill_manual(values = c("match" = "blue", "no match" = "red")) +  # Blue for match, Red for no match
  labs(x = "Coverage", y = "Frequency", title = "Histogram of Coverage by ZHeterozygositySexMatch") +
  theme_minimal() +
  theme(legend.title = element_blank(), legend.position = "top")

ggplot(SequencedIndividualsBirdIDsExtraDeduplicated, aes(x = Coverage, fill = RatioAutosomesZSexMatch)) +
  geom_histogram(bins = 30, alpha = 0.7, position = "identity", color = "black") +
  scale_fill_manual(values = c("match" = "blue", "no match" = "red")) +  # Blue for match, Red for no match
  labs(x = "Coverage", y = "Frequency", title = "Histogram of Coverage by RatioAutosomesZSexMatch") +
  theme_minimal() +
  theme(legend.title = element_blank(), legend.position = "top")

ggplot(na.omit(SequencedIndividualsBirdIDsExtraDeduplicated[, c("Coverage", "ZHeterozygositySexMatch")]), aes(x = Coverage, y = ZHeterozygositySexMatch)) +
  geom_point() +  # Adds scatter plot points
  theme_minimal() +  # Uses a clean, minimal theme
  labs(
    title = "Coverage vs. ZHeterozygositySexMatch",
    x = "Coverage",
    y = "ZHeterozygositySexMatch"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5),  # Center the title
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14)
  )
ggplot(na.omit(SequencedIndividualsBirdIDsExtraDeduplicated[, c("Coverage", "RatioAutosomesZSexMatch")]), aes(x = Coverage, y = RatioAutosomesZSexMatch)) +
  geom_point() +  # Adds scatter plot points
  theme_minimal() +  # Uses a clean, minimal theme
  labs(
    title = "Coverage vs. RatioAutosomesZSexMatch",
    x = "Coverage",
    y = "Ratio of Autosomes to Z-Sex Match"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5),  # Center the title
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14)
  )


SequencedIndividualsBirdIDsExtraDeduplicated <- SequencedIndividualsBirdIDsExtraDeduplicated %>%
  mutate(SexMatch = ifelse(ZHeterozygositySexMatch == "no match" & RatioAutosomesZSexMatch == "no match", "no match", "match"))

ggplot(na.omit(SequencedIndividualsBirdIDsExtraDeduplicated[, c("Coverage", "SexMatch")]), aes(x = Coverage, y = SexMatch)) +
  geom_point() +  # Adds scatter plot points
  theme_minimal() +  # Uses a clean, minimal theme
  labs(
    title = "Coverage vs. SexMatch",
    x = "Coverage",
    y = "Sex Match using heterozygosity and coverage"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5),  # Center the title
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14)
  )


ggplot(SequencedIndividualsBirdIDsExtraDeduplicated, aes(x = Coverage, fill = SexMatch)) +
  geom_histogram(bins = 30, alpha = 0.7, position = "identity", color = "black") +
  scale_fill_manual(values = c("match" = "blue", "no match" = "red")) +  # Blue for match, Red for no match
  labs(x = "Coverage", y = "Frequency", title = "Histogram of Coverage by SexMatching of both methods") +
  theme_minimal() +
  theme(legend.title = element_blank(), legend.position = "top")

PlinkSexCheck<- read.table("plink.sexcheck", header = TRUE, col.names=c("Filepath", "Filepath1", "PEDSEX", "SNPSEX", "STATUS","F"))

# Merge SequencedIndividualsBirdIDsExtraDeduplicated with PlinkSexCheck on Filepath, 
# but only keep rows where PlinkSexCheck$F is not NA or has values
SequencedIndividualsBirdIDsExtraDeduplicated <- SequencedIndividualsBirdIDsExtraDeduplicated %>%
  left_join(PlinkSexCheck %>% filter(!is.na(F)), by = "Filepath")


ggplot(SequencedIndividualsBirdIDsExtraDeduplicated, aes(x = F, fill = factor(Sex))) +
  geom_histogram(bins = 120, alpha = 0.7, position = "identity", color = "black") +
  scale_fill_manual(values = c("1" = "blue", "0" = "red")) +  # Blue for Sex=1, Red for Sex=0
  labs(x = "PLINK F", y = "Frequency", title = "Histogram of PLINK F with Sex-based Colors") +
  theme_minimal() +
  theme(legend.title = element_blank(), legend.position = "top")

SequencedIndividualsBirdIDsExtraDeduplicated <- SequencedIndividualsBirdIDsExtraDeduplicated %>%
    mutate(PLINKSex = case_when(
    F > 0.03 ~ "1",          
    F < 0.02 ~ "0",           
    #F >= 0.02 & F <= 0.03 ~ "Unsure"
    TRUE ~ NA_character_         # In case any other values exist, handle them
  ))
SequencedIndividualsBirdIDsExtraDeduplicated <- SequencedIndividualsBirdIDsExtraDeduplicated %>%
  mutate(PlinkSexMatch = ifelse(Sex == PLINKSex, "match", "no match"))



SequencedIndividualsBirdIDsExtraDeduplicated <- SequencedIndividualsBirdIDsExtraDeduplicated %>%
  mutate(SexMatch = ifelse(ZHeterozygositySexMatch == "no match" & RatioAutosomesZSexMatch == "no match" & PlinkSexMatch=="no match" , "no match", "match"))




```

## ROH

Include FROH as calculated by PLINK

```{r roh, eval=TRUE, warning=FALSE, include=TRUE }
##Large ROH (ROH>3.75Mb)
MySequencedIndividualsLargeROH<-read_table("largeROHnew.hom.indiv")
MySequencedIndividualsLargeROH$IID<-sub('.', '', MySequencedIndividualsLargeROH$IID)
MySequencedIndividualsLargeROH<-MySequencedIndividualsLargeROH %>% separate(IID, c('Filepath1', 'Filepath2', 'Filepath3', 'Plate','Filepath4','SeqID'), sep = '/', convert = TRUE)
MySequencedIndividualsLargeROH = subset(MySequencedIndividualsLargeROH, select = c(Plate, SeqID, NSEG, KB, KBAVG))
MySequencedIndividualsLargeROH$ID<-MySequencedIndividualsLargeROH$SeqID
MySequencedIndividualsLargeROH$FROH<-MySequencedIndividualsLargeROH$KB/1091184475*1000
names(MySequencedIndividualsLargeROH)[names(MySequencedIndividualsLargeROH) == 'FROH'] <- 'LargeFROH'

##Medium ROH (ROH>1.36Mb)
MySequencedIndividualsMediumROH<-read_table("mediumROHnew.hom.indiv")
MySequencedIndividualsMediumROH$IID<-sub('.', '', MySequencedIndividualsMediumROH$IID)
MySequencedIndividualsMediumROH<-MySequencedIndividualsMediumROH %>% separate(IID, c('Filepath1', 'Filepath2', 'Filepath3', 'Plate','Filepath4','SeqID'), sep = '/', convert = TRUE)
MySequencedIndividualsMediumROH = subset(MySequencedIndividualsMediumROH, select = c(Plate, SeqID, NSEG, KB, KBAVG))
MySequencedIndividualsMediumROH$ID<-MySequencedIndividualsMediumROH$SeqID
MySequencedIndividualsMediumROH$FROH<-MySequencedIndividualsMediumROH$KB/1091184475*1000
names(MySequencedIndividualsMediumROH)[names(MySequencedIndividualsMediumROH) == 'FROH'] <- 'MediumFROH'

##Small ROH (ROH>0.375Mb)
MySequencedIndividualsSmallROH<-read_table("smallROHnew.hom.indiv")
MySequencedIndividualsSmallROH$IID<-sub('.', '', MySequencedIndividualsSmallROH$IID)
MySequencedIndividualsSmallROH<-MySequencedIndividualsSmallROH %>% separate(IID, c('Filepath1', 'Filepath2', 'Filepath3', 'Plate','Filepath4','SeqID'), sep = '/', convert = TRUE)
MySequencedIndividualsSmallROH = subset(MySequencedIndividualsSmallROH, select = c(Plate, SeqID, NSEG, KB, KBAVG))
MySequencedIndividualsSmallROH$ID<-MySequencedIndividualsSmallROH$SeqID
MySequencedIndividualsSmallROH$FROH<-MySequencedIndividualsSmallROH$KB/1091184475*1000
names(MySequencedIndividualsSmallROH)[names(MySequencedIndividualsSmallROH) == 'FROH'] <- 'SmallFROH'

# Perform the joins
ROH <- MySequencedIndividualsLargeROH
ROH$MediumFROH<-MySequencedIndividualsMediumROH$MediumFROH
ROH$SmallFROH<-MySequencedIndividualsSmallROH$SmallFROH
ROH$LargeFROHbin<-ROH$LargeFROH
ROH$MediumFROHbin<-ROH$MediumFROH-(ROH$LargeFROH)
ROH$SmallFROHbin<-ROH$SmallFROH-(ROH$MediumFROH)

ROH <- ROH %>%
  arrange(SeqID, desc(Plate == "LIMSMERGED")) %>%
  distinct(SeqID, .keep_all = TRUE)

SequencedIndividualsBirdIDsExtraDeduplicated <- SequencedIndividualsBirdIDsExtraDeduplicated %>%
  left_join(ROH %>% select(SeqID, LargeFROHbin, MediumFROHbin, SmallFROHbin), by = "SeqID")


```

Call ROH using RZooRoH.
Sergio wants to use this in Zebra-finch, we can help each other out.

```{r eval=FALSE, echo=FALSE, include==FALSE}
#install.packages("RZooRoH", repos = "http://cran.us.r-project.org")
#Check out the package
library(RZooRoH)
## load data
SNPs <- zoodata(genofile = "rzoorohGP.gen", zformat = "gp")
## define model (e.g. if you simply want to capture all segments, you can put 4 classes 10, 100, 1000 + non-HBD)
mix4R <- zoomodel(K=4, base=10, err = 0.0025, layers=TRUE)
## run model on data
SNPsmodel <- zoorun(mix4R, SNPs, localhbd = TRUE)

```

## Life-history traits

```{r life-history, eval=TRUE, echo=FALSE, include=TRUE}
#Lifespan ----
##Read files
BirthDate <- read_csv("BirthDate27032023.csv", col_types = cols(BirthDate = col_date(format = "%d/%m/%Y"))) #In query table, this is BirdID
LastSeenYear <- read_csv("CurrentBTOextended27032023.csv") #In query table, this is CurrentBTOextended
BirdIDSexYear <- read_excel("BirdIDSexYear.xlsx")

##Make terms
###BirthYear from BirthDate
BirthDate <- BirthDate %>% 
  mutate(BirthYear = format(BirthDate, "%Y")) %>%
  mutate(BirthYear = as.numeric(BirthYear))

###Lifespan
Lifespan <- merge(BirthDate,LastSeenYear,by="BirdID", all = TRUE) %>% 
  mutate(LastSeenYea = as.numeric(LastSeenYea)) %>%
  mutate(BirthYear = as.numeric(BirthYear)) %>%
  mutate(Lifespan = LastSeenYea - BirthYear) %>%
  filter(LastSeenYea < 2022)  %>%
  select(BirdID,Lifespan,BirthYear,LastSeenYea)
colnames(Lifespan)[colnames(Lifespan) == 'LastSeenYea'] <- 'LastSeenYear'

##Link lifespan to dataframe
SequencedIndividualsBirdIDsExtraDeduplicated<- merge(SequencedIndividualsBirdIDsExtraDeduplicated, Lifespan, by="BirdID", all = TRUE)

##Link n offspring to dataframe
Offspring <- read_csv("Offspring27032023.csv",col_types = cols(BirthDate = col_date(format = "%d/%m/%Y")))
ROcount <- Offspring %>% filter(Confidence > 80) %>% count(Parent) %>% rename(BirdID = Parent,ReproductiveOutput = n)
SequencedIndividualsBirdIDsExtraDeduplicated<- merge(SequencedIndividualsBirdIDsExtraDeduplicated, ROcount, by="BirdID", all = TRUE)

SequencedIndividualsBirdIDsExtraDeduplicated <- SequencedIndividualsBirdIDsExtraDeduplicated %>%
  filter(!is.na(SeqID))


```

## Pedigree

Pedigree construction by *sequoia.* 572 SNPs were filtered for minor allele frequency \> 0.3, genotyping rate \> 99.9 and for linkage and also excluding SNPs on sex chromosomes and on chromosomes with \<90% imputation accuracy.

```{r pedigree, eval = TRUE, echo = TRUE, tidy=FALSE}
#Deduplicate samples for sequioa as it doesn't handle duplicates
DeduplicatedforSequoia<-data.frame(rep("0",length(SequencedIndividualsBirdIDsExtraDeduplicated$Filepath)),SequencedIndividualsBirdIDsExtraDeduplicated$Filepath)
write.table(DeduplicatedforSequoia, file = "DeduplicatedforSequoia.txt", sep = "\t",
            col.names = F, row.names = F, quote = FALSE)

#Rename samples for sequioa to BirdIDs
RenameforSequoia<-data.frame(rep("0",length(SequencedIndividualsBirdIDsExtraDeduplicated$Filepath)),SequencedIndividualsBirdIDsExtraDeduplicated$Filepath,rep("0",length(SequencedIndividualsBirdIDsExtraDeduplicated$Filepath)),SequencedIndividualsBirdIDsExtraDeduplicated$BirdID)
write.table(RenameforSequoia, file = "RenameforSequoia.txt", sep = "\t",
            col.names = F, row.names = F, quote = FALSE)

#Read in SNPs in .raw format
sw_GenoM_family <- GenoConvert(InFile = "reimputedinputfile_for_sequoia.raw", InFormat="raw")

#Read in pedigree from masterbayes
masterped <- read.csv("pedFINAL_var124567_combined_unique_20230413.csv")
head(masterped)
names(masterped)[names(masterped) == 'dad'] <- 'sire'
names(masterped)[names(masterped) == 'mum'] <- 'dam'

#drop unnecessary column
masterped <- masterped[-c(4)]

#Read in life history data
sw_LifeHistData <- read.csv("LifeHistoryData.csv", header = TRUE)
sw_LifeHistData<-sw_LifeHistData[!duplicated(sw_LifeHistData),]
sw_LifeHistData$Sex <- case_when(sw_LifeHistData$Sex == "0" ~ 1, 
                                 sw_LifeHistData$Sex == "1" ~ 2, 
                                 TRUE ~ 3)

#drop birds born before 1991 as they are not sampled
sw_LifeHistData <- sw_LifeHistData[sw_LifeHistData$Year > 1990,]

#drop birds not in cousin
sw_LifeHistData <- sw_LifeHistData[sw_LifeHistData$Island == "CN",]

#drop unnecessary column
sw_LifeHistData <- sw_LifeHistData[,-2]
sw_LifeHistData <- sw_LifeHistData[,-3]

sw_LifeHistData<-na.omit(sw_LifeHistData)

#Looks like lots of duplicate BirdIDs so create new life hist file
BirthDate <- read_csv("BirthDate27032023.csv", col_types = cols(BirthDate = col_date(format = "%d/%m/%Y"))) #In query table, this is BirdID
BirdIDSexYear <- read_excel("BirdIDSexYear.xlsx")
lifehist<- merge(BirthDate, BirdIDSexYear, by= "BirdID", all= TRUE)%>%
  select(BirdID,Sex,BirthDate)
lifehist<-lifehist[!duplicated(lifehist),]
lifehist$Sex <- case_when(lifehist$Sex == "0" ~ 1, 
                          lifehist$Sex == "1" ~ 2, 
                                 TRUE ~ 3)
lifehist$BirthDate<-format(as.Date(lifehist$BirthDate, format="%Y/%m/%d"),"%Y")
colnames(lifehist)[colnames(lifehist) == 'BirthDate'] <- 'Year'
lifehist<-lifehist[-c(1, 2), ]  

# 1. Initial diagnostics --------------------------------------------------

# Check genotype calls across the dataset (-9 = missing calls)
summary.factor(as.factor(sw_GenoM_family))/(nrow(sw_GenoM_family)*ncol(sw_GenoM_family))

# Check for Mendelian errors in the data at the locus level
GenoM_checks <- SnpStats(sw_GenoM_family, Ped = masterped[,c("id", "dam", "sire")])

ggplot(GenoM_checks, aes(x = OHdam)) + geom_histogram() + theme_classic()
ggplot(GenoM_checks, aes(x = OHsire)) + geom_histogram() + theme_classic()
sw_OHLLR_DATABASE <- CalcOHLLR(Pedigree = masterped[,c("id", "dam", "sire")], GenoM = sw_GenoM_family)
ggplot(sw_OHLLR_DATABASE, aes(y = LLRdam, x = OHdam)) + geom_point(alpha = 0.6) + theme_classic()
ggplot(sw_OHLLR_DATABASE, aes(y = LLRsire, x = OHsire)) + geom_point(alpha = 0.6) + theme_classic()
SummarySeq(sw_OHLLR_DATABASE, Panels = "OH")

# 2. Run basic sequoia ------------------------------------

# Note modules used to build up from basic input check to analysis:
# i.   'pre': Input check
# ii.  'dup': check for duplicates
# iii. 'par': parentage assignment
# iv.  'ped': full pedigree reconstruction

# Just parentage assignment (Module = "par"). 
sw_family_sequoia_justparents <- sequoia(GenoM = sw_GenoM_family, LifeHistData = lifehist, Module = "par", quiet="verbose", args.AP = list(MaxAgeParent = c(19)))

# A 94.0% match to the masterbayes pedigree
sw_family_pedigree_justparents_comparison <- PedCompare(masterped[,c("id", "dam", "sire")], sw_family_sequoia_justparents$Pedigree, Plot=TRUE, Symmetrical = TRUE)

SummarySeq(SeqList = sw_family_sequoia_justparents)


# Full pedigree reconstruction (Module = "ped"). Hashed out of knit because takes too long.
#sw_family_sequoia <- sequoia(GenoM = sw_GenoM_family, LifeHistData = lifehist, Module = "ped", quiet="verbose", args.AP = list(MaxAgeParent = c(19)))
#sw_family_pedigree_comparison <- PedCompare(masterped[,c("id", "dam", "sire")], sw_family_sequoia$Pedigree, Plot=TRUE)$MergedPed

```

## Demographic history

Demographic history inference using GONE.

```{r demographic_history, eval = TRUE, echo = TRUE, tidy=FALSE}
#Generation time calculation
##add parent birth date from "BirthDate" to "Offspring"
BirthDate$Parent<-BirthDate$BirdID
OffspringParentBirthDate<-merge(Offspring, BirthDate, by= "Parent", all =TRUE)
OffspringParentBirthDate$ParentBirthDate<-OffspringParentBirthDate$BirthDate.y

##calculate mean birth date difference
OffspringParentBirthDate$BirthDateDifference<-OffspringParentBirthDate$BirthDate.x-OffspringParentBirthDate$ParentBirthDate
BirthDateDifference<-(!is.na (OffspringParentBirthDate$BirthDateDifference))
BirthDateDifference <- subset(OffspringParentBirthDate$BirthDateDifference, !is.na(OffspringParentBirthDate$BirthDateDifference))

mean(OffspringParentBirthDate$BirthDateDifference, na.rm = TRUE)
1764/365 ##4.8 generation time

#Load output from GONE
GONE<-read.table("outfileLD_Ne_estimates",header=F, fill=TRUE)
GONE <- GONE[-c(1, 2), ]
rownames(GONE) <- NULL
colnames(GONE) <- c("Generation", "Ne")
GONE$YearsAgo<-as.numeric(GONE$Generation)*4.8
GONE$Ne<-as.numeric(GONE$Ne)
GONE$Generation<-as.numeric(GONE$Generation)


#Plot. NB I will need to iterate this to get confidence intervals
ggplot(GONE, aes(x = Generation, y = Ne)) +
  geom_point(color = "blue") +
  scale_x_continuous(breaks = seq(0, 2000, by = 200)) +  # Set specific breaks
  labs(x = "Generation", y = "Ne", title = "Plot of Generations vs Ne") +
  theme_minimal()
```
## Inbreeding depression

Investigate co-variation between individual inbreeding coefficients and key life-history traits. It looks like there is more evidence of inbreeding depression from recent inbreeding. Include control variables once Pinto returns from field, as agreed.

```{r inbreeding_depression, eval = TRUE, echo = TRUE}
##Inbreeding (FROH) depression (Lifespan and N offspring produced in lifetime)
LargeROHxLifespan<-ggplot(dplyr::filter(SequencedIndividualsBirdIDsExtraDeduplicated), aes(x=LargeFROHbin, y=Lifespan)) + geom_point() + geom_smooth(method='lm') 
LargeROHxLifespan

MediumROHxLifespan<-ggplot(dplyr::filter(SequencedIndividualsBirdIDsExtraDeduplicated), aes(x=MediumFROHbin, y=Lifespan)) + geom_point() + geom_smooth(method='lm') 
MediumROHxLifespan

SmallROHxLifespan<-ggplot(dplyr::filter(SequencedIndividualsBirdIDsExtraDeduplicated), aes(x=SmallFROHbin, y=Lifespan)) + geom_point() + geom_smooth(method='lm') 
SmallROHxLifespan


```

## Inbreeding avoidance

Calculate relatedness of all pedigree pairs confirmed by sequoia and how that matches up to pairwise relatedness of all adults.
Sen and Sergio have done some work here, we can help each other out.

```{r inbreeding_avoidance, eval = TRUE, echo = TRUE}


# Ensure the necessary columns are of the same type (character)
Lifespan$BirdID <- as.character(Lifespan$BirdID)


# Merge Rel.both with Lifespan for IID1 and IID2
Rel.both.alive <- Rel.both %>%
  left_join(Lifespan, by = c("IID1" = "BirdID"), suffix = c("_1", "_"))
Rel.both.alive <- Rel.both.alive %>%
  left_join(Lifespan, by = c("IID2" = "BirdID"), suffix = c("_1", "_2"))

# Check for overlap in BirthYear and LastSeenYear
Rel.both.alive <- Rel.both.alive %>%
  filter(BirthYear_1 <= LastSeenYear_2 & BirthYear_2 <= LastSeenYear_1)

masterped_nonmissing <- masterped[!is.na(masterped$mum) & !is.na(masterped$dad), ]

# Ensure both are data frames
masterped_nonmissing <- as.data.frame(masterped_nonmissing)
Rel.both.alive <- as.data.frame(Rel.both.alive)

# Create a helper column to represent the two possible pairings
masterped_nonmissing$pair1 <- paste(masterped_nonmissing$mum, masterped_nonmissing$dad, sep = "_")
masterped_nonmissing$pair2 <- paste(masterped_nonmissing$dad, masterped_nonmissing$mum, sep = "_")

# Create a new column in Rel.both.alive for the pair combination of IID1 and IID2
Rel.both.alive$pair <- paste(Rel.both.alive$IID1, Rel.both.alive$IID2, sep = "_")

# Subset Rel.both.alive to include only rows where the pair is present in masterped_nonmissing
Rel.both.alive.partners <- Rel.both.alive %>%
  filter(pair %in% masterped_nonmissing$pair1 | pair %in% masterped_nonmissing$pair2)


# Calculate the mean of Rel.both.alive.partners$R.GRM
mean_R_GRM <- mean(Rel.both.alive.partners$R.GRM, na.rm = TRUE)

# Create a ggplot2 histogram for Rel.both.alive$R.GRM with the mean line
ggplot(Rel.both.alive, aes(x = R.GRM)) +
  geom_histogram(binwidth = 0.1, fill = "lightblue", color = "black", alpha = 0.7) +
  geom_vline(xintercept = mean_R_GRM, color = "red", linetype = "dashed", size = 1.5) +
  labs(title = "Histogram of R.GRM", x = "R.GRM", y = "Frequency") +
  theme_minimal()







```

## GWAS

In progress. Adapting script from Stoffel et. al 2021: <https://github.com/mastoffel/sheep_ID/blob/master/6_alt_gwas_annual_survival_bothA_sep.R>
We need to fit an inbreeding depression model for each SNP.
Then use the functional annotations if any SNP turns up.
Chuen has done some  work on this with microbiome/fitness, we can help each other out.

```{r gwas, eval = FALSE, echo = TRUE}

library(lme4)
library(tidyverse)
library(broom.mixed)
library(snpStats)
library(data.table)
library(furrr)

# for running on server, this allows to run an array job for splitting
# the GWAS up in parts
part_inp  <- commandArgs(trailingOnly=TRUE)
if (!(length(part_inp) == 0)) {
        part <- as.numeric(part_inp[[1]])
} else {
        # if no part selected, take first 1000
        part <- 1000
}


# GRM PCs from plink
pcs <- read_delim("data/ann_surv_pca.txt", " ", col_names = TRUE) %>% 
        mutate(id = as.character(id))

# roh data
file_path <- "data/roh.hom"
roh_lengths <- fread(file_path)

# plink name
sheep_plink_name <- "data/sheep_geno_imputed_oar_filt"
# read merged plink data
sheep_bed <- paste0(sheep_plink_name, ".bed")
sheep_bim <- paste0(sheep_plink_name, ".bim")
sheep_fam <- paste0(sheep_plink_name, ".fam")
full_sample <- read.plink(sheep_bed, sheep_bim, sheep_fam)

# make list with all parts
all_snps <- 1:nrow(full_sample$map)
all_parts <- split(all_snps, ceiling(seq_along(all_snps )/1000)) # every part runs 500 models
snp_indices <- all_parts[[part]]

# filter map data
snps_map_sub <- as_tibble(full_sample$map[snp_indices, ])
# additive genotypes
geno_sub <- as_tibble(as(full_sample$genotypes[, snps_map_sub$snp.name], Class = "numeric"),
                      rownames = "id")
# survival data
# survival data preprocessing
annual_survival <- fitness_data %>% 
        # filter na rows
        filter_at(vars(survival, froh_all, birth_year, sheep_year), ~ !is.na(.)) %>% 
        mutate(age_cent = age - mean(age, na.rm = TRUE),
               age_cent2 = age_cent^2,
               age_std = as.numeric(scale(age)),
               age_std2 = age_std^2,
               # times 10 to estimate a 10% percent increase
               froh_all10 = froh_all * 10,
               froh_all10_cent = froh_all10 - mean(froh_all10, na.rm = TRUE),
               lamb = ifelse(age == 0, 1, 0),
               lamb_cent = lamb - mean(lamb, na.rm = TRUE),
               lamb = as.factor(lamb)) %>% 
        as.data.frame() 

#roh_lengths <- as.data.table(roh_lengths)
# check whether snp is in ROH for a given individual

setkey(roh_lengths, IID)
roh_id_per_snp <- function(i) {
        position <- as.numeric(snps_map_sub[i, "position"])
        chromosome <- as.numeric(snps_map_sub[i, "chromosome"])
        # varname <- paste0("roh", i)
        #roh <- as.numeric((roh_lengths$POS1 <= position) & (roh_lengths$POS2 >= position) & (roh_lengths$CHR == chromosome))
        #roh_lengths$roh <- roh
        roh_lengths[, roh := as.numeric((CHR == chromosome) & (POS1 <= position) & (POS2 >= position))]
        #roh_lengths[, roh := fifelse((POS1 <= position)&(POS2 >= position)&(CHR == chromosome), 1, 0)]
        roh_id <- roh_lengths[,  .(roh = max(roh)), by = c("IID")]$roh
}

roh_ind <- map(1:nrow(snps_map_sub), roh_id_per_snp)
roh_df <- as.data.frame(do.call(cbind, roh_ind))
names(roh_df) <- paste0("roh_", snps_map_sub$snp.name)
roh_df$id <- as.character(unique(roh_lengths$IID))


# make some space
rm(full_sample)

# which chromosomes do the snps span?
chrs <- unique(snps_map_sub$chromosome)
froh_no_chr <- paste0("froh_no_chr", chrs)
# join additive and roh data to survival for gwas
annual_survival_gwas <- annual_survival %>% 
        #mutate_at(vars(starts_with("froh_no_chr")), scale) %>% 
        dplyr::select(id, survival, sex, twin, lamb, birth_year, sheep_year, mum_id, age_std, age_std2, {{ froh_no_chr }}) %>% 
        left_join(pcs, by = "id") %>% 
        left_join(geno_sub, by = "id") %>% 
        left_join(roh_df, by = "id") %>% 
        as_tibble()

snp_names <- snps_map_sub$snp.name
#snp_names <- top_snps %>% group_by(chromosome) %>% top_n(-2, p.value) %>% .$snp.name
for (i in snp_names) {
        # dummy coding
        annual_survival_gwas[[paste0("roh_0_", i)]] <- as.numeric((annual_survival_gwas[[i]] == 0) & (annual_survival_gwas[[paste0("roh_", i)]] == 1))
        annual_survival_gwas[[paste0("roh_2_", i)]] <- as.numeric((annual_survival_gwas[[i]] == 2) & (annual_survival_gwas[[paste0("roh_", i)]] == 1))
        annual_survival_gwas[[paste0("roh_", i)]] <- NULL
}

# time saver function for modeling
nlopt <- function(par, fn, lower, upper, control) {
        .nloptr <<- res <- nloptr(par, fn, lb = lower, ub = upper, 
                                  opts = list(algorithm = "NLOPT_LN_BOBYQA", print_level = 1,
                                              maxeval = 1000, xtol_abs = 1e-6, ftol_abs = 1e-6))
        list(par = res$solution,
             fval = res$objective,
             conv = if (res$status > 0) 0 else res$status,
             message = res$message
        )
}

# focal SNP, chromosome of focal snp, data
run_gwas <- function(snp, data) {
        # for mean froh without focal chr
        chr <- as.numeric(snps_map_sub[snps_map_sub$snp.name == snp, "chromosome"])
        froh_no_chr <- paste0("froh_no_chr", chr)
        
        formula_snp <- as.formula(paste0("survival ~ 1 + sex + twin + age_std + age_std2 + ", 
                                         froh_no_chr, " + ",
                                         "pc1 + pc2 + pc3 + pc4 + pc5 + pc6 + pc7 + ",
                                         #"pc1 + pc2 + pc3 + pc4 +",
                                         snp, "+ ", paste0("roh_0_", snp), "+",  paste0("roh_2_", snp), "+ (1|birth_year) + (1|sheep_year) + (1|id)"))
        #snp, "+ ", paste0("roh_", snp), " + (1|sheep_year) + (1|id)"))
        mod <- glmer(formula = formula_snp,
                     data = data, family = "binomial",
                     control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))
        out <- broom.mixed::tidy(mod)
        out
}

safe_run_gwas <- purrr::safely(run_gwas)

# 

snps_sub <- snps_map_sub$snp.name
# split into pieces of 50 SNPs 
num_parts <- round(length(seq_along(snps_sub )) / 50)
snps_pieces <- split(snps_sub, cut(seq_along(snps_sub), num_parts, labels = FALSE))
roh_pieces_0 <- map(snps_pieces, function(x) paste0("roh_0_", x)) #### to change here
roh_pieces_2 <- map(snps_pieces, function(x) paste0("roh_2_", x)) 

annual_survival_gwas_pieces <- 
        pmap(list(snps_pieces, roh_pieces_0, roh_pieces_2), function(snps_piece, roh_piece_0, roh_piece_2) {
                annual_survival_gwas %>% dplyr::select(id:pc7, one_of(c(snps_piece, roh_piece_0, roh_piece_2 )))   
        })

# clean up
rm(annual_survival, annual_survival_gwas, fitness_data, geno_sub, 
   roh_lengths, roh_pieces_0, roh_pieces_2, sheep_ped, roh_df)

# set up plan
plan(multiprocess, workers = 6)

# increase maxSize
options(future.globals.maxSize = 3000 * 1024^2)

all_out <- future_map2(snps_pieces, annual_survival_gwas_pieces, function(snps, data) {
        out <- purrr::map(snps, safe_run_gwas, data)
})

all_out_simple <- purrr::flatten(all_out)
saveRDS(all_out_simple, file = paste0("output/GWAS_roh_sep_part", "_", part, ".rds"))


### testing
# all_out <- future_map2(snps_pieces[1], annual_survival_gwas_pieces[1], function(snps, data) {
#         out <- purrr::map(snps[1:3], safe_run_gwas, data)
# })
```


##Sample mismatches

Investigating sex mismatches between field database and genotype data and duplicate pruning.

```{r sample_mismatch, eval=TRUE, echo=FALSE, include=TRUE}
library(dplyr)
#I used three genomic methods of sexing: Z heterozygosity, Autosome:Z coverage and PLINK’s --sex-check function. There are 29/1977 samples where all three methods agreed on a mismatch to the field pedigree.
SexMismatched <- SequencedIndividualsBirdIDsExtraDeduplicated %>%
  filter(SexMatch == "no match" &
         (Het < 0.076 | Het > 0.084) &
         (RatioAutosomesZ < 1.98 | RatioAutosomesZ > 2.02) &
         (F < 0.021 | F > 0.029))

#27/29 of sex mismatches are also pedigree mismatches. These birds need investigation.
#This could be because 1) we have the same sample but have incorrectly sexed the bird in the field database, or 2) we have sampled a different bird for sequencing.

PedigreeMismatch<-sw_family_pedigree_justparents_comparison[["Mismatch"]]

SexMismatched$BirdID<-as.character(SexMismatched$BirdID)
SexMismatched$PedigreeMismatch <- apply(SexMismatched, 1, function(row) {
  any(row["BirdID"] %in% unlist(PedigreeMismatch))
})

#There are 2/29 sex mismatched birds but are pedigree matched. Both birds had no offsping.
#For the BirdID: 6141, SeqID: 20_TAGGCCTCCT-GTGGTCAATT_L001__all_mapped_rehead.bam, coverage is too low to ascertain sex. INCLUDE AS F?
#For BirdID: 1668, SeqID: 98-1015_221007_L003__all_mapped_rehead.bam, the bird was only seen twice in 2002 (Unknown) and 2003 (SEEN1). Field database is marked as male by molecular sexing but female by observation. INCLUDE AS F?


openxlsx::write.xlsx(SexMismatched, file ="SexMismatched.xlsx", quote=FALSE)

##Genomic relatedness comparison to masterbayes pedigree relatedness

#Convert masterbayes pedigree to relatedness

Rped.old<-CalcRped(masterped, OUT="DF")
Rel.snp<- read.table("plink2.grm", stringsAsFactors = FALSE)
Rel.id <- read.table("plink2.grm.id", stringsAsFactors=FALSE)
Rel.id<-merge(Rel.id, RenameforSequoia, by.x='V2', by.y='SequencedIndividualsBirdIDsExtraDeduplicated.Filepath') %>% select(V1,SequencedIndividualsBirdIDsExtraDeduplicated.BirdID)
Rel.snp[,1] <- as.character(factor(Rel.snp[,1], labels=Rel.id[,2]))
Rel.snp[,2] <- as.character(factor(Rel.snp[,2], labels=Rel.id[,2]))
names(Rel.snp) <- c("IID1", "IID2", "nSNPS", "R.GRM")
Rel.snp <- Rel.snp[Rel.snp$IID1 != Rel.snp$IID2,] 
Rel.both <- merge(data.table(Rel.snp[,c(1,2,4)], key=c("IID1", "IID2")),
                  data.table(Rped.old, key=c("IID1", "IID2")), all.x=TRUE)
rm(Rel.snp, Rped.old)   # clean up: remove large dataframes
round(cor(Rel.both[, c("R.GRM","R.ped")],use="pairwise.complete"), 3)
hexbin::hexbinplot(Rel.both$R.GRM ~ Rel.both$R.ped,
                   xbins=100, aspect=1, 
                   xlim=c(0.05,1), ylim=c(-.7, 2.8),
                   xlab="Pedigree relatedness", ylab="Genomic relatedness",
                   trans=log10, inv=function(x) 10^x,
                   colorcut=seq(0,1,length=14), maxcnt=10^6.5,
                   colramp = function(n) {grDevices::hcl.colors(n, palette='Berlin')})

hb <- hexbin::hexbin(Rel.both$R.GRM ~ Rel.both$R.ped,
                     xbins=100, xbnds=c(-.05, 2), ybnds=c(-.7, 2.8),
                     xlab="Pedigree relatedness", ylab="Genomic relatedness")
hbp <- hexbin::plot(hb,
                                  trans=log10, inv=function(x) 10^x,
                                  colorcut=seq(0,1,length=14), maxcnt=10^6.5,
                                  colramp = function(n) {grDevices::hcl.colors(n, palette='Berlin')}
                                  )
hexbin::hexVP.abline(hbp$plot.vp, lm(Rel.both$R.GRM ~ Rel.both$R.ped))

#We can try find which bird it is we incorrectly sampled for sequencing.
##Change sex of samples to "unknown" in sequoia and re-run parentage assignment. Then see which bird is it's parents and offspring and compare to microsat pedigree.
lifehistsexunknown<-lifehist
lifehistsexunknown$Sex[lifehistsexunknown$BirdID %in% SexMismatched$BirdID] <- 3

# Just parentage assignment (Module = "par"). 
sw_family_sequoia_justparents <- sequoia(GenoM = sw_GenoM_family, LifeHistData = lifehistsexunknown, Module = "par", quiet="verbose", args.AP = list(MaxAgeParent = c(19)))

sw_family_pedigree_justparents_comparison <- PedCompare(masterped[,c("id", "dam", "sire")], sw_family_sequoia_justparents$Pedigree, Plot=TRUE, Symmetrical = TRUE)

# Subset the genomic relatedness pairwise matrix for sex mismatched birds that have genomic relatedness > 0.8 as these birds may have been sequenced twice unintentionally.
Rel.sexmismatched.both <- Rel.both[(
  (Rel.both$IID1 %in% SexMismatched$BirdID | Rel.both$IID2 %in% SexMismatched$BirdID) &
  (Rel.both$R.GRM > 0.8)
), ]

# Using genomic relatedness >0.75 does not always indicate a duplicate. It is possible that a pair could have a high genomic relatedness (e.g. >0.75) and a low pedigree relatedness (< 0.1) if the pedigree was gappy. For example, two birds  with high relatedness (perhaps full sibs with high inbreeding coefficients) with unknown parents. Birds with unknown parents are treated as founders in a pedigree, regardless of how early/late in the long-term study they were born/hatched. 
# Therefore, load a duplicate check datafile using PLINK, calling duplicates by --king-table-filter 0.35 for 487,107 SNPs (genotyping rate > 99.9 and MAF > 0.3)
UnintentionalDuplicates<- read.table("unintentionalduplicates.kin0", stringsAsFactors = FALSE)

# Loop through rows of UnintentionalDuplicates and add BirdID from SequencedIndividualsBirdIDsExtraDeduplicated to V1 where V2 matches Filepath
# Loop through rows of UnintentionalDuplicates and add BirdID from SequencedIndividualsBirdIDsExtraDeduplicated to V1 where V2 matches Filepath
#Add BirdIDs
UnintentionalDuplicates$V1[UnintentionalDuplicates$V2 %in% SequencedIndividualsBirdIDsExtraDeduplicated$Filepath] <- 
  SequencedIndividualsBirdIDsExtraDeduplicated$BirdID[match(UnintentionalDuplicates$V2, SequencedIndividualsBirdIDsExtraDeduplicated$Filepath)]

UnintentionalDuplicates$V3[UnintentionalDuplicates$V4 %in% SequencedIndividualsBirdIDsExtraDeduplicated$Filepath] <- 
  SequencedIndividualsBirdIDsExtraDeduplicated$BirdID[match(UnintentionalDuplicates$V4, SequencedIndividualsBirdIDsExtraDeduplicated$Filepath)]

# Rename columns in UnintentionalDuplicates
colnames(UnintentionalDuplicates) <- c("BirdID1", "Filepath1", "BirdID2", "Filepath2", "NSNPs", "V6", "V7", "Kinship")

# Subset UnintentionalDuplicates to remove rows where Kinship is NA
UnintentionalDuplicates <- UnintentionalDuplicates[!is.na(UnintentionalDuplicates$Kinship), ]

# Initialize the new column with FALSE (or NA) in UnintentionalDuplicates
UnintentionalDuplicates$SexMismatchedDuplicate <- FALSE

# Loop through each BirdID in SexMismatched to check if it matches in either BirdID1 or BirdID2 in UnintentionalDuplicates
for (i in 1:nrow(SexMismatched)) {
  # Find rows in UnintentionalDuplicates where the BirdID is present in either BirdID1 or BirdID2
  matching_rows <- UnintentionalDuplicates[UnintentionalDuplicates$BirdID1 == SexMismatched$BirdID[i] |
                                           UnintentionalDuplicates$BirdID2 == SexMismatched$BirdID[i], ]
  
  # If a match is found, set SexMismatchedDuplicate to TRUE for the corresponding rows
  if (nrow(matching_rows) > 0) {
    UnintentionalDuplicates$SexMismatchedDuplicate[UnintentionalDuplicates$BirdID1 == SexMismatched$BirdID[i] |
                                                  UnintentionalDuplicates$BirdID2 == SexMismatched$BirdID[i]] <- TRUE
  }
}


# Create a new column in SexMismatched to store the Kinship values
SexMismatched$Kinship <- NA  # Initialize with NA

# Loop through each BirdID in SexMismatched to check if it matches in either BirdID1 or BirdID2 in UnintentionalDuplicates
for (i in 1:nrow(SexMismatched)) {
  # Find rows in UnintentionalDuplicates where the BirdID is present in either BirdID1 or BirdID2
  matched_row <- UnintentionalDuplicates[UnintentionalDuplicates$BirdID1 == SexMismatched$BirdID[i] |
                                        UnintentionalDuplicates$BirdID2 == SexMismatched$BirdID[i], ]
  
  # If a match is found, assign the corresponding Kinship value
  if (nrow(matched_row) > 0) {
    SexMismatched$Kinship[i] <- matched_row$Kinship[1]  # Assuming one match per BirdID
  }
}



#Any duplicate samples  in sequoia input SNPs? Yes, 29.

DuplicatesSequoia<-sw_family_sequoia_justparents[["DupGenotype"]]

# Initialize the new column in UnintentionalDuplicates with FALSE (or NA)
UnintentionalDuplicates$SequoiaDuplicate <- FALSE

UnintentionalDuplicates$BirdID1 <- as.character(UnintentionalDuplicates$BirdID1)
UnintentionalDuplicates$BirdID2 <- as.character(UnintentionalDuplicates$BirdID2)
DuplicatesSequoia$ID1 <- as.character(DuplicatesSequoia$ID1)
DuplicatesSequoia$ID2 <- as.character(DuplicatesSequoia$ID2)

# Perform a left join to merge UnintentionalDuplicates with relevant columns from DuplicatesSequoia
UnintentionalDuplicates <- UnintentionalDuplicates %>%
  left_join(DuplicatesSequoia %>%
              select(ID1, ID2, Mismatch, SnpdBoth, LLR), 
            by = c("BirdID1" = "ID2", "BirdID2" = "ID1")) %>%
  mutate(SequoiaDuplicate = ifelse(BirdID1 %in% DuplicatesSequoia$ID1 | 
                                     BirdID1 %in% DuplicatesSequoia$ID2 |
                                     BirdID2 %in% DuplicatesSequoia$ID1 | 
                                     BirdID2 %in% DuplicatesSequoia$ID2, 
                                   TRUE, FALSE))


#Subset for PLINK Kinship > 0.496 as there are no duplicates over this value 
UnintentionalDuplicates <- subset(UnintentionalDuplicates,UnintentionalDuplicates$Kinship > 0.496)

#There are 29 duplicates as found by sequoia and confirmed with PLINK. Now we need to investigate which is the correct BirdID out of two. 
#11 can be confirmed using sex
#329 is in fact 1176
#478 is in fact 3444
#4882 is in fact 3643
#4883 is in fact 1607
#5542 is in fact 6136
#5546 is in fact 6175
#5550 is in fact 4961
#5585 is in fact 4943
#6150 is in fact 5786
#6862 is in fact 6670
#1276 is in fact 1765

#15 can be confirmed by comparing with masyerbayes mother
# Perform a left join to look up the mother (dam) from masterped using id
UnintentionalDuplicates$BirdID1 <- as.character(UnintentionalDuplicates$BirdID1)
UnintentionalDuplicates$BirdID2 <- as.character(UnintentionalDuplicates$BirdID2)
Rel.both$IID1<-as.character(Rel.both$IID1)
Rel.both$IID2<-as.character(Rel.both$IID2)

masterped$id <- as.character(masterped$id)

UnintentionalDuplicates <- UnintentionalDuplicates %>%
  left_join(masterped %>%
              select(id, dam), 
            by = c("BirdID1" = "id")) %>%
  rename(BirdID1Mum = dam)

UnintentionalDuplicates <- UnintentionalDuplicates %>%
  left_join(masterped %>%
              select(id, dam), 
            by = c("BirdID2" = "id")) %>%
  rename(BirdID2Mum = dam)

UnintentionalDuplicates$BirdID1Mum <- as.character(UnintentionalDuplicates$BirdID1Mum)
UnintentionalDuplicates$BirdID2Mum <- as.character(UnintentionalDuplicates$BirdID2Mum)

# Step 1: Perform the joins for BirdID1 and BirdID1Mum
UnintentionalDuplicates <- UnintentionalDuplicates %>%
  left_join(Rel.both %>%
              select(IID1, IID2, R.GRM) %>%
              rename(R.GRM_1 = R.GRM), 
            by = c("BirdID1" = "IID1", "BirdID1Mum" = "IID2")) %>%
  left_join(Rel.both %>%
              select(IID1, IID2, R.GRM) %>%
              rename(R.GRM_2 = R.GRM), 
            by = c("BirdID1" = "IID2", "BirdID1Mum" = "IID1")) %>%
  mutate(BirdID1MumRel = coalesce(R.GRM_1, R.GRM_2))

# Step 2: Perform the joins for BirdID2 and BirdID2Mum
UnintentionalDuplicates <- UnintentionalDuplicates %>%
  left_join(Rel.both %>%
              select(IID1, IID2, R.GRM) %>%
              rename(R.GRM_3 = R.GRM), 
            by = c("BirdID2" = "IID1", "BirdID2Mum" = "IID2")) %>%
  left_join(Rel.both %>%
              select(IID1, IID2, R.GRM) %>%
              rename(R.GRM_4 = R.GRM), 
            by = c("BirdID2" = "IID2", "BirdID2Mum" = "IID1")) %>%
  mutate(BirdID2MumRel = coalesce(R.GRM_3, R.GRM_4))

PedigreeComparison<-sw_family_pedigree_justparents_comparison[["MergedPed"]]

# Join for BirdID1 and create BirdID1MumMatch
UnintentionalDuplicates <- UnintentionalDuplicates %>%
  left_join(PedigreeComparison %>% select(id, dam.class), 
            by = c("BirdID1" = "id")) %>%
  mutate(BirdID1MumMatch = dam.class) %>%
  select(-dam.class)  # Remove dam.class column if not needed

# Join for BirdID2 and create BirdID2MumMatch
UnintentionalDuplicates <- UnintentionalDuplicates %>%
  left_join(PedigreeComparison %>% select(id, dam.class), 
            by = c("BirdID2" = "id")) %>%
  mutate(BirdID2MumMatch = dam.class) %>%
  select(-dam.class)  # Remove dam.class column if not needed

# Remove the unwanted columns R.GRM_1, R.GRM_2, R.GRM_3, and R.GRM_4 from UnintentionalDuplicates
UnintentionalDuplicates <- UnintentionalDuplicates %>%
  select(-R.GRM_1, -R.GRM_2, -R.GRM_3, -R.GRM_4)

#3 can be confirmed comparing with offspring.
#1218/1182. 1182 is probably 1218 because 1218 matches with three different offspring in masterbayes and sequoia
#498/1303. 1303 is probably 498 because 498 has 0.5 relatedness with one offspring (but not the other).
#864/3871. 3871 is probably 864 because 864 matches with offspring.



#Create a BirdIDConsensus column for what the BirdID of the duplicate pair should be
UnintentionalDuplicates <- UnintentionalDuplicates %>%
  mutate(BirdIDConsensus = case_when(
    BirdID1 %in% SexMismatched$BirdID ~ as.character(BirdID2),
    BirdID2 %in% SexMismatched$BirdID ~ as.character(BirdID1),
    BirdID1MumMatch == "Match" ~ as.character(BirdID1),
    BirdID2MumMatch == "Match" ~ as.character(BirdID2),
    BirdID1MumRel > 0.5 ~ as.character(BirdID1),
    BirdID2MumRel > 0.5 ~ as.character(BirdID2),
    BirdID1 == 1218 | BirdID2 == 1218 ~ "1218",
    BirdID1 == 498  | BirdID2 == 498  ~ "498",
    BirdID1 == 864  | BirdID2 == 864  ~ "864",
    TRUE ~ NA_character_
  ))

UnintentionalDuplicates <- UnintentionalDuplicates %>%
  mutate(BirdIDMissing = case_when(
    BirdID1 == BirdIDConsensus ~ BirdID2,  # If BirdID1 equals BirdIDConsensus, set BirdIDMissing to BirdID2
    BirdID2 == BirdIDConsensus ~ BirdID1,  # If BirdID2 equals BirdIDConsensus, set BirdIDMissing to BirdID1
    TRUE ~ NA_character_                       # If neither condition is met, set BirdIDMissing to NA (numeric NA)
  ))

openxlsx::write.xlsx(UnintentionalDuplicates, file ="UnintentionalDuplicates.xlsx", quote=FALSE)
getwd()

#Re-run sequoia removing the duplicates and only picking the higher coverage sample
UnintentionalDuplicates <- UnintentionalDuplicates %>%
  mutate(
    BirdID1Coverage = case_when(
      BirdID1 %in% SequencedIndividualsBirdIDsExtraDeduplicated$BirdID ~ 
        SequencedIndividualsBirdIDsExtraDeduplicated$Coverage[match(BirdID1, SequencedIndividualsBirdIDsExtraDeduplicated$BirdID)],
      TRUE ~ NA_real_  # If no match, set to NA
    ),
    BirdID2Coverage = case_when(
      BirdID2 %in% SequencedIndividualsBirdIDsExtraDeduplicated$BirdID ~ 
        SequencedIndividualsBirdIDsExtraDeduplicated$Coverage[match(BirdID2, SequencedIndividualsBirdIDsExtraDeduplicated$BirdID)],
      TRUE ~ NA_real_  # If no match, set to NA
    )
  )

UnintentionalDuplicates <- UnintentionalDuplicates %>%
  mutate(
    BirdIDtoExclude = case_when(
      BirdID1Coverage > BirdID2Coverage ~ BirdID2,  # If BirdID1Coverage is greater, set BirdIDtoExclude to BirdID2
      BirdID1Coverage < BirdID2Coverage ~ BirdID1,  # If BirdID1Coverage is less, set BirdIDtoExclude to BirdID1
      TRUE ~ NA_character_  # If both are equal or other cases, set BirdIDtoExclude to NA
    )
  )


# Ensure BirdIDtoExclude is a character vector for comparison
exclude_values <- as.character(UnintentionalDuplicates$BirdIDtoExclude)

# Check if sw_GenoM_family has row names
if (!is.null(rownames(sw_GenoM_family))) {
  
  # Get the row names of sw_GenoM_family
  row_names <- rownames(sw_GenoM_family)
  
  # Find the values in row_names that are in exclude_values
  rows_to_exclude <- row_names %in% exclude_values
  
  # Exclude the rows from sw_GenoM_family where the row names match exclude_values
  sw_GenoM_family_filtered <- sw_GenoM_family[!rows_to_exclude, , drop = FALSE]
  
  print("Rows have been filtered successfully.")
} else {
  print("sw_GenoM_family does not have row names.")
}

sw_dedup_family_sequoia_justparents <- sequoia(GenoM = sw_GenoM_family_filtered, LifeHistData = lifehist, Module = "par", quiet="verbose", args.AP = list(MaxAgeParent = c(19)))

# A 95.1% match to the masterbayes pedigree (1% increase after deduplicating)
sw_dedup_family_pedigree_justparents_comparison <- PedCompare(masterped[,c("id", "dam", "sire")], sw_dedup_family_sequoia_justparents$Pedigree, Plot=TRUE, Symmetrical = TRUE)

SummarySeq(SeqList = sw_family_sequoia_justparents)

#Why are there duplicates?
view(Identifiers)

Identifiers$BirdID<-as.character(SheffieldSubmissions$BirdID)

#It looks like Plate 53 is responsible for 12 duplicates!
UnintentionalDuplicatesMoreInfo<-UnintentionalDuplicates
for (i in 1:nrow(UnintentionalDuplicatesMoreInfo)) {
  
  # Match BirdID1 with Identifiers$BirdID and return Identifiers$Plate in BirdID1Plate
  match1 <- which(Identifiers$BirdID == UnintentionalDuplicatesMoreInfo$BirdID1[i])
  if (length(match1) > 0) {
    UnintentionalDuplicatesMoreInfo$BirdID1Plate[i] <- Identifiers$Plate[match1]
  }
  
  # Match BirdID2 with Identifiers$BirdID and return Identifiers$Plate in BirdID2Plate
  match2 <- which(Identifiers$BirdID == UnintentionalDuplicatesMoreInfo$BirdID2[i])
  if (length(match2) > 0) {
    UnintentionalDuplicatesMoreInfo$BirdID2Plate[i] <- Identifiers$Plate[match2]
  }
}

for (i in 1:nrow(UnintentionalDuplicatesMoreInfo)) {
  
  # Match BirdID1 with Identifiers$BirdID and return Identifiers$Plate in BirdID1Plate
  match1 <- which(Identifiers$BirdID == UnintentionalDuplicatesMoreInfo$BirdID1[i])
  if (length(match1) > 0) {
    UnintentionalDuplicatesMoreInfo$BirdID1BloodID[i] <- Identifiers$BloodID[match1]
  }
  
  # Match BirdID2 with Identifiers$BirdID and return Identifiers$Plate in BirdID2Plate
  match2 <- which(Identifiers$BirdID == UnintentionalDuplicatesMoreInfo$BirdID2[i])
  if (length(match2) > 0) {
    UnintentionalDuplicatesMoreInfo$BirdID2BloodID[i] <- Identifiers$BloodID[match2]
  }
}

for (i in 1:nrow(UnintentionalDuplicatesMoreInfo)) {
  
  # Match BirdID1 with Identifiers$BirdID and return Identifiers$Plate in BirdID1Plate
  match1 <- which(Identifiers$BirdID == UnintentionalDuplicatesMoreInfo$BirdID1[i])
  if (length(match1) > 0) {
    UnintentionalDuplicatesMoreInfo$BirdID1BloodTubeNumber[i] <- Identifiers$BTN[match1]
  }
  
  # Match BirdID2 with Identifiers$BirdID and return Identifiers$Plate in BirdID2Plate
  match2 <- which(Identifiers$BirdID == UnintentionalDuplicatesMoreInfo$BirdID2[i])
  if (length(match2) > 0) {
    UnintentionalDuplicatesMoreInfo$BirdID2BloodTubeNumber[i] <- Identifiers$BTN[match2]
  }
}

for (i in 1:nrow(UnintentionalDuplicatesMoreInfo)) {
  
  # Match BirdID1 with Identifiers$BirdID and return Identifiers$Plate in BirdID1Plate
  match1 <- which(Identifiers$BirdID == UnintentionalDuplicatesMoreInfo$BirdID1[i])
  if (length(match1) > 0) {
    UnintentionalDuplicatesMoreInfo$BirdID1Box[i] <- Identifiers$Box[match1]
  }
  
  # Match BirdID2 with Identifiers$BirdID and return Identifiers$Plate in BirdID2Plate
  match2 <- which(Identifiers$BirdID == UnintentionalDuplicatesMoreInfo$BirdID2[i])
  if (length(match2) > 0) {
    UnintentionalDuplicatesMoreInfo$BirdID2Box[i] <- Identifiers$Box[match2]
  }
}

for (i in 1:nrow(UnintentionalDuplicatesMoreInfo)) {
  
  # Match BirdID1 with Identifiers$BirdID and return Identifiers$Plate in BirdID1Plate
  match1 <- which(Identifiers$BirdID == UnintentionalDuplicatesMoreInfo$BirdID1[i])
  if (length(match1) > 0) {
    UnintentionalDuplicatesMoreInfo$BirdID1Plate_number[i] <- Identifiers$`Plate Number`[match1]
  }
  
  # Match BirdID2 with Identifiers$BirdID and return Identifiers$Plate in BirdID2Plate
  match2 <- which(Identifiers$BirdID == UnintentionalDuplicatesMoreInfo$BirdID2[i])
  if (length(match2) > 0) {
    UnintentionalDuplicatesMoreInfo$BirdID2Plate_number[i] <- Identifiers$`Plate Number`[match2]
  }
}

#There are 50 samples in Plate 53
sum(Identifiers$Plate == 53, na.rm = TRUE)
#Is Plate 53 also implicated in the sex mismatches? Yes, 12/26 sex-mismatched samples are from Plate 53.
for (i in 1:nrow(SexMismatched)) {
  
  match <- which(Identifiers$BirdID == SexMismatched$BirdID[i])
  
  if (length(match) > 0) {
    SexMismatched$PlateSheffield[i] <- Identifiers$Plate[match]
  }
}

#Of the sequoia mismatches, how many are from Plate 53? 42/156!
# Assuming PedigreeMismatch and Identifiers are data frames
for (i in 1:nrow(PedigreeMismatch)) {
  
  # Match PedigreeMismatch$id with Identifiers$BirdID
  match_id <- which(Identifiers$BirdID == PedigreeMismatch$id[i])
  if (length(match_id) == 1) {  # Ensure exactly one match
    PedigreeMismatch$idPlate[i] <- Identifiers$Plate[match_id]
  } else {
    PedigreeMismatch$idPlate[i] <- NA  # If no match or multiple matches, assign NA
  }
  
  # Match PedigreeMismatch$dam.1 with Identifiers$BirdID
  match_dam <- which(Identifiers$BirdID == PedigreeMismatch$dam.1[i])
  if (length(match_dam) == 1) {  # Ensure exactly one match
    PedigreeMismatch$dam.1Plate[i] <- Identifiers$Plate[match_dam]
  } else {
    PedigreeMismatch$dam.1Plate[i] <- NA  # If no match or multiple matches, assign NA
  }
  
  # Match PedigreeMismatch$sire.1 with Identifiers$BirdID
  match_sire <- which(Identifiers$BirdID == PedigreeMismatch$sire.1[i])
  if (length(match_sire) == 1) {  # Ensure exactly one match
    PedigreeMismatch$sire.1Plate[i] <- Identifiers$Plate[match_sire]
  } else {
    PedigreeMismatch$sire.1Plate[i] <- NA  # If no match or multiple matches, assign NA
  }
}
count <- sum(
  grepl("53", PedigreeMismatch$idPlate) |
  grepl("53", PedigreeMismatch$dam.1Plate) |
  grepl("53", PedigreeMismatch$sire.1Plate)
)
count

#Explore how Plate53 looks visually. I don't see any pattern here.
IdentifiersPlate53<-subset(Identifiers, Identifiers$Plate == 53)
# Assuming IdentifiersPlate53 is your data frame with columns BirdID and well

# Create an empty matrix to store the BirdID values
result_matrix <- matrix(NA, nrow = 12, ncol = 8)

# Set the row and column names (A to H for columns, 1 to 12 for rows)
colnames(result_matrix) <- LETTERS[1:8]
rownames(result_matrix) <- 1:12

# Loop through each row in IdentifiersPlate53 to assign BirdID to the appropriate place in the matrix
for (i in 1:nrow(IdentifiersPlate53)) {
  # Extract the well location (e.g., "A1", "B2")
  well <- IdentifiersPlate53$well[i]
  
  # Extract column letter (A to H)
  col <- substr(well, 1, 1)
  
  # Extract row number (1 to 12)
  row <- as.integer(substr(well, 2, 2))
  
  # If the row number is a two-digit number (e.g., "A10"), adjust the extraction for the row
  if (nchar(well) == 3) {
    row <- as.integer(substr(well, 2, 3))
  }
  
  # Convert column letter to column number (A = 1, B = 2, ..., H = 8)
  col_num <- which(LETTERS[1:8] == col)
  
  # Assign the BirdID to the appropriate place in the result matrix
  result_matrix[row, col_num] <- IdentifiersPlate53$BirdID[i]
}

# Print the result matrix
result_matrix


#Which sex mismatched samples remain to be resolved?
SexMismatchedToResolve <- SexMismatched %>%
  mutate(DuplicateSample = BirdID %in% c(UnintentionalDuplicates$BirdID1, UnintentionalDuplicates$BirdID2))

#11/29 can be resolved as duplicates
#3/29 genomic sex can't reliably be estimated as coverage < 0.32, include as microsatellite sex?
#7/29 I think can be resolved by investigating relationships with close relatives:
  #3088 (M, but geno F) never had offspring (parents only are mismatched)- check how it was sexed in database. Has 2 blood samples. Mother is #1716 and father is #1774, which is the masterbayes parents of 5145. SWAP WITH 5145.

  #5145 (F, but geno M) never had offspring (parents only are mismatched)- check how it was sexed in database. Has 2 blood samples. Mother is #1429 and father is #1886, which is the masterbayes parents of 3088. SWAP WITH 3088.

  #6156 (M, but geno F) is  #6160. SWAP.

  #6160 (M, but geno F) is  #6156. SWAP.

  #6166 mismatched parents and is replaced in sequoia by #5538 as father to #6766 and #6862. Recorded as BrM 5 years in a row from 2015-2019. Has 5 blood samples. This is probably the real #6862 as #5538 is also father to #6670 who is also #6862. SWAP?

  #774 (F, but geno M) never has offspring according to sequoia, and gets replaced by #329 (#1176) as father to #1176. Has 2 blood samples. BrF from 1996-2002. This is probably the real #329? SWAP?

#8/29 I am not sure how to treat:

  #1229 (M, but geno F) mother is #825 and then inbreeds with #825 to have  an offspring #1537.

  #1435 (M, but geno F) never had offspring according to sequoia, but had three offspring according to masterbayes. Has 4 blood samples and BrM from 2002-2004. Father is #1181. #1435 is related to 870, 1181 and 1052 by > 0.4. 

  #1676 (M, but geno F) never had offspring according to sequoia, but had one offspring according to masterbayes and gets replaced as father by #1695 (who is BrM in neighboring territory 46 vs 45). Has 1 blood sample as ABX. Mother is #1390. SEX CONFLICT IN DATABASE, INCLUDE AS F?

  #5554 (M, but geno F) mismatched parents and also is a sequoia mother to #5773 and #6034, both times replacing #5111. Has 3 blood samples. Mother is #4863 and father is #1762. Related to 5634, 1884, 511, 5553, 1762 > 0.4. Gets assigned female in pre-deduplicated sequoia pedigree and gets assigned male in deduplicated sequoia pedigree.

  #5555 (M, but geno F) never had offspring (parents only are mismatched)- check how it was sexed in database. Has 1 blood sample as FL. Mother is #3500 and father is #1762. #5554 and #5555 are both males so unlikely to have been swapped. Related to 3500 and 1762 > 0.4. 5555 is 5512, 5512 is 6151, 6151 is 6178, 6178 is 6176 (6176 is a missing sample)?

  #5827 (F, but geno M) never had offspring (mother only is mismatched)- check how it was sexed in database. Has 1 blood sample as FL. Mother is #3086. Related to 3086, 5339, 5598 > 0.35.

  #6049 (M, but geno F) never had offspring (parents only are mismatched)- check how it was sexed in database. Has 3 blood samples. Father is #1750.

  #6149 (M, but geno F) never had offspring (parents only are mismatched)- check how it was sexed in database. Has 3 blood samples. Mother is #1707 and father is #5822.


# Subset the SexMismatched data frame where DuplicateSample is FALSE
SexMismatched <- SexMismatched %>%
  filter(DuplicateSample == FALSE)

SexMismatched <- SexMismatched %>%
  mutate(BirdIDConsensus = case_when(
    BirdID == 774  ~ 329,   # If BirdID is 774, set BirdIDConsensus to 329
    BirdID == 1617 ~ 1617,  # If BirdID is 3088, set BirdIDConsensus to 5145
    BirdID == 3088 ~ 5145,  # If BirdID is 3088, set BirdIDConsensus to 5145
    BirdID == 5145 ~ 3088,  # If BirdID is 3088, set BirdIDConsensus to 5145
    BirdID == 5833 ~ 5833,  # If BirdID is 5833, set BirdIDConsensus to 5833
    BirdID == 6156 ~ 6160,  # If BirdID is 6156, set BirdIDConsensus to 6160
    BirdID == 6160 ~ 6156,  # If BirdID is 6160, set BirdIDConsensus to 6156
    BirdID == 6166 ~ 6862,  # If BirdID is 6166, set BirdIDConsensus to 6862
    BirdID == 6141 ~ 6141,  # If BirdID is 6141, set BirdIDConsensus to 6141
    TRUE ~ NA_real_         # If no condition is met, set BirdIDConsensus to NA (numeric NA)
  ))

# Where did the SexMismatched samples come from?
# Sex mismatched seem to be evenly distributed along timescale ordered by BloodID
ggplot(SequencedIndividualsBirdIDsExtraDeduplicated %>% filter(Coverage > 0.32), 
       aes(x = ID, y = SexMatch)) +
  geom_point(na.rm = TRUE) +  # Scatter plot, ignoring NA values
  labs(x = "BloodID", y = "SexMatch", title = "BloodID vs SexMatch (Coverage > 0.32)") +
  theme_minimal()


Seychelles_warbler_traits<-SequencedIndividualsBirdIDsExtraDeduplicated

```



